What Drives State-Sponsored Violence?: Evidence from Extreme Bounds Analysis and Ensemble Learning Models

### 2018-11-30 paper-killings/manuscript.tex
upload

To overcome these limitations and provide a better understanding of government atrocity, I employ extreme bounds analysis and random forests to identify the most robust determinants of state-sponsored atrocities. My approach is similar to Hegre and Sambanis' \citeyear{hegre2006sensitivity} seminal analysis on the causes of civil war onset, but like \citet{bell2015examining}, \citep{hill
} and \citet{jones2018there}, I provide additional tests to verify whether complex interactions and nonlinearities are driving the statistical results. In conducting this analysis, I address three debates in the mass violence literature:

@@

More broadly, the goal of this study is to examine whether the quantitative scholarship is able to identify the main explanators of mass atrocities and, if positive, which explanators are more robust relative to others \citep[2]{bell2015examining}. My analysis tests the sensitivity of 40 variables on a sample of 177 countries from 1945 to 2013. My findings partially confirm previous research -- poor, unstable countries are more likely to witness the regime employ atrocity \citep[e.g.,][]{goldsmith2013forecasting,harff2003no,krain1997state}. However, many of the factors scholars often cite as observable indicators of such instability -- regime transitions, coups d'état, the presence of militias, etc. -- are not good proxies for instability. Thus, policymakers may be looking for the incorrect signs of impending atrocity when seeking to prevent its onset. Furthermore, I find that the conclusions scholars draw regarding the likelihood of government atrocity largely depend on whether they combine peace and war years or just analyse periods of civil wars, as patterns in mass killings differ dramatically across these contexts. Lastly, I find that genocide and politicide follow vastly different patterns of onset than other forms of state-sponsored mass murder. This is further evidence that different logics govern different forms of political violence \citep{stanton2013terrorism}.

@@

@@

@@

>>>>

To overcome these limitations and provide a better understanding of government atrocity, I employ extreme bounds analysis and random forests to identify the most robust determinants of state-sponsored atrocities. My approach is similar to Hegre and Sambanis' \citeyear{hegre2006sensitivity} seminal analysis on the causes of civil war onset, but like \citet{bell2015examining}, \citep{hill2014empirical} and \citet{jones2018there}, I provide additional tests to verify whether complex interactions and nonlinearities are driving the statistical results. More broadly, my goal is to examine whether the quantitative scholarship is able to identify robust explanators of mass atrocities and, if positive, how they compare to each other in explaining power.

While recent studies have progressively adopted stronger identification strategies, the majority of the literature on mass killings is comprised of cross-country regressions. Large-$n$ analyses have benefits that are often overlooked. For instance, quantitative studies allow scholars to assess how generalisable are the results of specific mechanisms. These studies provide a safeguard against the perils of sampling on the dependent variable, and they increase the external validity of micro-level statistical or qualitative findings \citep{bell2015examining,king1994designing}. Additionally, cross-national comparisons show how structural factors condition immediate causes of mass violence. Atrocities are multicausal phenomena, and large-$n$ studies can point to important interactions that scholars would otherwise miss. In this specific study, there is the additional advantage of running all the models with the same dependent variable, what constitutes an effective method of replication of the original findings.

In conducting this analysis, I address three debates in the mass violence literature:

@@

My analysis tests the sensitivity of 40 variables on a sample of 177 countries from 1945 to 2013. My findings partially confirm previous research -- poor, unstable countries are more likely to witness the regime employ atrocity \citep[e.g.,][]{goldsmith2013forecasting,harff2003no,krain1997state}. However, many of the factors scholars often cite as observable indicators of such instability -- regime transitions, coups d'état, the presence of militias, etc. -- are not good proxies for instability. Thus, policymakers may be looking for the incorrect signs of impending atrocity when seeking to prevent its onset. Furthermore, I find that the conclusions scholars draw regarding the likelihood of government atrocity largely depend on whether they combine peace and war years or just analyse periods of civil wars, as patterns in mass killings differ dramatically across these contexts. Lastly, I find that genocide and politicide follow vastly different patterns of onset than other forms of state-sponsored mass murder. This is further evidence that different logics govern different forms of political violence \citep{stanton2013terrorism}.

@@

\subsection{Main Model}%
\label{sub:main_model}

@@

\subsection{Mass Killings during Civil Wars}%
\label{sub:mass_killings_during_civil_wars}

@@

\subsection{Mass Killings in and after the Cold War}%
\label{sub:mass_killings_in_and_after_the_cold_war}

\subsection{Genocides and Politicides}%
\label{sub:genocides_and_politicides}
